---
title: "Combining oly assessments"
format: html
editor_options: 
  chunk_output_type: console
---

## About

This script is used to clean and standardize individual oly population assessments, so they can be combined into one main file used in PSRF's oly work webpage.

## Setup

```{r}
# ===== Load libraries ===== #
library(here)
library(tidyverse)
library(janitor)
library(lubridate)

# ===== Read in data ===== #
# Dyes Inlet sites ----

# Chico Bay
chicobay_2019_raw <- read_csv(here("data", "raw", "chicobay_2019-08-30_counts_raw.csv")) |> clean_names()

chicobay_2021_raw <- read_csv(here("data", "raw", "chicobay_2021-04-29_counts_raw.csv")) |> clean_names()

# Oyster Bay
oysterbay_2020_raw <- read_csv(here("data", "raw", "oysterbay_2020-07-03_counts_raw.csv")) |> clean_names()

# Silverdale
silverdale_raw <- read_csv(here("data", "raw", "silverdale_oyster_assessment_countsHH.csv")) |>
  clean_names() # previously processed by MK & HH

# Mud Bay
mudbay_2014_raw <- read_csv(here("data", "raw", "mudbay_2014-06-26_counts_raw.csv")) |>
  clean_names()

# Fidalgo Bay sites ----

fidalgo_2016_raw <- read_csv(here("data", "raw", "fidalgobay_2016_counts_raw.csv"))

weaverlingspit_2020_raw <- read_csv(here("data", "raw", "weaverlingspit_2020_counts_raw.csv"), na = c("ND", "NA")) |>
  clean_names()

fidalgo_2023_raw <- read_csv(here("data", "raw", "fidalgobay_2023_counts_raw.csv"))

# 

```

## Approach

Each larger waterbody will get their own cleaned .csv file. These files will include these columns:

- **waterbody**: The larger waterbody
- **site**: The standardized site name
- **recorded_site**: The site name as recorded by the original surveyors
- **date**: yyyy-mm-dd
- **latitude**: Location of sample (when recorded)
- **longitude**: Location of sample (when recorded)
- **survey_area**: Project or Reference. Project if survey contains any project footprint, can include area outside of project footprint
- **sample_id**: Unique sample ID. References to spatial data when applicable
- **sample_method**: random point, haphazard point, transect
- **sample_unit_m2**: Size of quadrat used
- **live_oly_count**: Number of live olympia oysters counted in this sample
- **dead_oly_count**: Number of dead olympia oysters counted in this sample
- **notes**: Notes!

After each survey sheet looks like this, I will combine them to the larger waterbody.

Once combined, I will calculate density in a new data frame.

```{r}
# Save list of columns I want to keep
column_names <- c("waterbody", "site", "recorded_site", "date", "latitude", "longitude",
                  "survey_area", "sample_id", "sample_method", "sample_unit_m2",
                  "live_oly_count", "dead_oly_count", "notes")
```


## Dyes Inlet Sites

### Chico Bay

Chico bay had 500 seeded cultch bags and 500 cubic yards of shell spread on tidelands covering 5 acres in 2018

#### Chico Bay 2019

MK looked briefly for site map / spatial data in drive and ArcGIS Online and did not find (3/4/26)

```{r}
# Rename columns ----
chicobay_2019 <- chicobay_2019_raw |>
  rename("survey_area" = "survey_area_type",
         "sample_unit_m2" = "unit_m2",
         "live_oly_count" = "live_count",
         "dead_oly_count" = "dead_count")

# Standardize survey area nomenclature ----
chicobay_2019 <- chicobay_2019 |>
  mutate(survey_area = 
           case_when(survey_area == "reference - 0.5MLLW" ~ "Reference",
                     survey_area == "enhancement2018" ~ "Project"))

# Add columns needed ----
chicobay_2019 <- chicobay_2019 |>
  mutate(recorded_site = "Chico Bay",
         latitude = NA_real_,
         longitude = NA_real_,
         sample_method = "Random")

# Keep and order columns based on pre-made list ----
chicobay_2019 <- chicobay_2019 |>
  select(all_of(column_names))

# Fix data types ----
chicobay_2019 <- chicobay_2019 |>
  mutate(dead_oly_count = NA_real_,
         notes = NA_character_)
```

All set

#### Chico Bay 2021

Spatial data in ArcGIS : https://psrf.maps.arcgis.com/apps/mapviewer/index.html?layers=7d42c910eb284452ac521e38e3cfcf89

3/3/26 - struggling to download attribute layer to easily grab lat and long. Thinking it's because I am not the owner and the organization has it's settings so that I can't download the data. Ask (Alex? Kari?) about this soon! 

```{r}
# Rename columns ----
chicobay_2021 <- chicobay_2021_raw |>
  rename("survey_area" = "survey_area_type",
         "sample_unit_m2" = "unit_m2",
         "live_oly_count" = "live_count",
         "dead_oly_count" = "dead_count")

# Add columns needed ----
chicobay_2021 <- chicobay_2021 |>
  mutate(recorded_site = "Chico Bay",
         latitude = NA_real_,
         longitude = NA_real_,
         sample_method = "Random")

# Keep and order columns based on pre-made list ----
chicobay_2021 <- chicobay_2021 |>
  select(all_of(column_names))

# Standardize survey area nomenclature ----
chicobay_2021 <- chicobay_2021 |>
  mutate(survey_area = 
           case_when(survey_area == "enhancement2018" ~ "Project"))

# Fix data types ----
chicobay_2021 <- chicobay_2021 |>
  mutate(dead_oly_count = NA_real_)
```

Good! 

Combine Chico Bay files:

```{r}
chico_bay <- rbind(chicobay_2019, chicobay_2021)
```

```{r}
# Clean up environment
rm(list = setdiff(ls(), c("chico_bay", "column_names")))
```

Would like to confirm with someone that we only visited Chico bay in 2019 and 2021!

### Oyster Bay

Brian says that this survey includes are within the footprint of a 2011 bulk shell project, but also a large area of the surrounding tideland, effectively sampling all "productive" areas.

Lat and long is included in this data. Here is the ArcGIS link anyways: https://psrf.maps.arcgis.com/apps/mapviewer/index.html?webmap=04b1c80881b7496e8bd6fec2e81b833b

```{r}
# Rename columns ----
oysterbay_2020 <- oysterbay_2020_raw |>
  rename("live_oly_count" = "live_count",
         "sample_unit_m2" = "sample_unit_m")

# Add columns needed ----
oysterbay_2020 <- oysterbay_2020 |>
  mutate(recorded_site = "Oyster Bay",
         sample_method = "Haphazard",
         dead_oly_count = NA_real_,
         survey_area = "Project")   # If area includes a project at all, = project, even if there is area outside sampled

# Keep and order columns based on pre-made list ----
oysterbay_2020 <- oysterbay_2020 |>
  select(all_of(column_names))

# Standardize site names ----
oysterbay_2020 <- oysterbay_2020 |>
  mutate(waterbody = "Dyes Inlet",
         site = "Oyster Bay")

# Fix data types ----
oysterbay_2020 <- oysterbay_2020 |>
  mutate(notes = NA_character_)
```

Nice

**Other Oyster Bay Notes**

PSRF staff visited Oyster bay in July 2012 to look at the emergent shell area from the 2011 shell enhancement. Though the data mapping the perimeter of the shell enhancement was lost, a note mentions that the enhancement is "located as intended", abutting the western edge of the natural oyster bed.

### Mud Bay

```{r}
# Rename columns ----
mudbay_2014 <- mudbay_2014_raw |>
  rename("survey_area" = "survey_area_type",
         "sample_method" = "survey_method",
         "sample_unit_m2" = "unit_m2",
         "live_oly_count" = "live_count")

# Add columns needed ----
mudbay_2014 <- mudbay_2014 |>
  mutate(recorded_site = "Mud Bay",
         dead_oly_count = NA_real_,
         latitude = NA_real_,
         longitude = NA_real_,
         notes = NA_character_)

# Standardize survey area and method capitalization ----
mudbay_2014 <- mudbay_2014 |>
  mutate(survey_area = str_to_title(survey_area),
         sample_method = str_to_title(sample_method))

# Keep and order columns based on pre-made list ----
mudbay_2014 <- mudbay_2014 |>
  select(all_of(column_names))

# Data types look good
```

Nice! 

Did not find anything else for Mud Bay when searching the drive.

### Silverdale

```{r}
# Rename columns ----
silverdale <- silverdale_raw |>
  rename("recorded_site" = "location",
         "survey_area" = "pop_type",
         "sample_method" = "method",
         "sample_unit_m2" = "total_sample_area_m2")

# Add columns needed ----
silverdale <- silverdale |>
  mutate(waterbody = "Dyes Inlet")

# Keep and order columns based on pre-made list ----
silverdale <- silverdale |>
  select(all_of(column_names))

# Standardize capitalization and nomenclature within columns ----
silverdale <- silverdale |>
  mutate(survey_area = case_when(survey_area == "ref" ~ "Reference",
                                 survey_area == "project" ~ "Project",
                                 .default = survey_area),
         sample_method = str_to_sentence(sample_method))

# Fix data types ----
silverdale <- silverdale |>
  mutate(latitude = NA_real_,
         longitude = NA_real_)
```

Lot's of silverdale stuff in ArcGIS. Need to go digging for this at some point. Perhaps this is something that Alex can help drum up.

### Combining Dyes

As of 3/3/2026, this is all of the data I am aware of for oyster count assessments for Dyes Inlet.

```{r}
# Combine sites for Dyes Inlet waterbody into one file and save to use later ----
dyes_inlet <- rbind(chico_bay, oysterbay_2020, mudbay_2014, silverdale)

write_csv(dyes_inlet, here("data", "processed", "dyes_inlet_oly_counts.csv"))
```


## Fidalgo Bay Sites

### 2016

Note: this survey measured plots near but only outside of the shell enhancement plots. Going to call survey_area reference for now

```{r}
# Rename columns ----
fidalgo_2014 <- fidalgo_2016_raw |>
  # new name = old name
  rename("site" = "accepted_site",
         "sample_method" = "method",
         "live_oly_count" = "live_count",
         "dead_oly_count" = "dead_count",
         "sample_id" = "sampleID") |>
  # Create survey_area column ----
  mutate(survey_area = "reference")

# Keep and order columns based on pre-made list ----
fidalgo_2014 <- fidalgo_2014 |>
  select(all_of(column_names))
```

### Weaverling Spit 2019

We have some information here - but I need some help interpreting before including


### Weaverling Spit 2020

```{r}
# Add and rename columns ---- 
fidalgo_2020 <- weaverlingspit_2020_raw |>
  # new name = old name
  rename("recorded_site" = "site",
         "sampleunit_m2" = "unit_m2",
         "live_oly_count" = "live_count",
         "dead_oly_count" = "dead_count") |>
  mutate(survey_area = "project",
         sample_method = "transect",
         site = "Fidalgo Bay west",
         latitude = NA_real_,
         longitude = NA_real_)

# Keep and order columns based on list ----
fidalgo_2020 <- fidalgo_2020 |>
  select(all_of(column_names))
```


### 2023

```{r}
# Add and rename columns ----
fidalgo_2023 <- fidalgo_2023_raw |>
  rename("recorded_site" = "site",
         "sample_method" = "method",
         "sample_id" = "sampleID",
         "live_oly_count" = "livesamplecount") |>
  mutate(survey_area = NA_character_,
         site = case_when(recorded_site == "March's Point" ~ "Fidalgo Bay east",
                          recorded_site == "Causeway - North" ~ "Fidalgo Bay west",
                          recorded_site == "Causeway - South" ~ "Fidalgo Bay west",
                          recorded_site == "Weaverling Spit" ~ "Fidalgo Bay west",
                          recorded_site == "Southern Sloughs" ~ "Fidalgo Sloughs",
                          recorded_site == "East Slough" ~ "Fidalgo Sloughs",
                          recorded_site == "Crandle Spit (MRC Area 13)" ~ "Fidalgo Bay east",
                          recorded_site == "Weaverling Spit - High Puddle" ~ "Fidalgo Bay west",
                          .default = recorded_site),
         dead_oly_count = NA_real_)

# Keep and order columns based on list
fidalgo_2023 <- fidalgo_2023 |>
  select(all_of(column_names))
```


### Combine Fidalgo

```{r}
fidalgo <- rbind(fidalgo_2014, fidalgo_2020, fidalgo_2023)
```

```{r}
# Clean environment (remove all except fidalgo and column_names)
rm(list = setdiff(ls(), c("fidalgo", "column_names")))
```


```{r}
# What unique site / survey_area pairs are there? 
fidalgo |>
  distinct(site, survey_area)
```

I think all of those NAs are from the 2023 survey, when I am unsure about which points are in project footprints and which are not. Hoping to figure out which is which using these maps in GIS: https://psrf.maps.arcgis.com/apps/mapviewer/index.html?layers=7925bc1ae8c14452940d12e83480a464

For now, I am going to ignore that (knowing fully that will probably need to be addressed)

```{r}
# Create month / year columns ----
fidalgo <- fidalgo |>
  mutate(month = month(date),
         year = year(date))

# Make sampleunit_m2 numeric
fidalgo <- fidalgo |>
  mutate(sampleunit_m2 = as.numeric(sampleunit_m2))

# Need to logic through different quadrat sizes when I have more brainpower

```

